### 2019-02-06 03b_R-data.Rmd
Added package data summary

---
title: "Data in R packages"
author: "Karthik"
date: "2/6/2019"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE}
suppressPackageStartupMessages(library(tidyverse))
df <- suppressMessages(read_csv("package_files_with_metadata.csv"))
data_folder <-  df %>% filter(location == "data") %>% pull(package) %>% unique %>% length
extdata_folder   <- df %>% filter(location == "extdata") %>% pull(package) %>% unique %>% length
packages_with_data <- df %>% filter(grepl("data|extdata", location)) %>% pull(package) %>% unique %>% length
unique_packages <- df %>% pull(package) %>% unique %>% length

#  Do some summaries
only_data_extdata <- df %>% filter(grepl("data|extdata", location))
types <- only_data_extdata %>% count(extension) %>% arrange(desc(n)) %>% mutate(percentage = n/nrow(only_data_extdata))

percent <- function(x, digits = 2, format = "f", ...) {
  paste0(formatC(100 * x, format = format, digits = digits, ...), "%")
}

```

Of the 15539 packages on Central R Archive Network (CRAN),  `r packages_with_data` contain datasets either as binary data (`r data_folder` packages) or as external datasets (`r extdata_folder`). Binary files comprise a bulk of the data shipped in the `data` folder (`r types[1:2, ]$percentage %>% sum %>% percent`) with other plain text formats such as txt, csv, dat, json comprising less than one percent of data shipped in packages.