### 2021-08-11 response-letter/response-letter.Rmd
minor changes to get the table to print (trick was to include `column_spec(1)`)

---
title: "Response to reviewer comments"
authors: Nicholas Tierney, Karthik Ram
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libs, include = FALSE}
library(dplyr)
library(googlesheets4)
library(knitr)
library(kableExtra)
```

```{r fix-long-url}

new_comment_line <- "The authors need to further clarify how their recommendations align with - and differ from - the slew of other recommendations for data management which exist (in articles - e.g https://doi.org/10.1371/journal.pcbi.1003542, and at the level of institutional/regional data stewardship - e.g. from the Research Data Alliance, or from libraries: https://www.monash.edu/library/researchers /n/researchdata/guidelines/sharing.) I think the key to differentiating this work is again making the community/audience explicit from the very beginning"

```

```{r read-sheet, include = FALSE}
our_table_id <- gs4_get(
  "1fIl8jy72XWD4liFZVlx6WvsnXp7ld8fs7eZvwVJpSy0"
  )

our_table <- read_sheet(ss = our_table_id,
                        .name_repair = janitor::make_clean_names) %>% 
  select(ID = id, 
         comment = reviewer_comment, 
         response) %>% 
  mutate(comment = 
           case_when(
             ID == "R1-2" ~ new_comment_line,
             TRUE ~ comment
           )) 
```

Dear Editor in chief of Patterns,

Thank you for considering our commentary article “Incentives are the biggest barriers to widespread data sharing” for publication in Patterns.

We thank the reviewers for time and comments. We have addressed comments from the reviewers and believe the paper is clearer and better as a result. Our response to reviewer comments are presented in the table below. The "ID" column value "R1-1" refers to reviewer 1 comment 1, and so on.

Regards,

Nicholas Tierney & Karthik Ram

```{r show-table}
our_table %>% 
  kbl(booktabs = TRUE,
      longtable = TRUE) %>% 
  kable_styling(latex_options = c("striped",
                                  "hold_position", 
                                  "repeat_header"),
                full_width = TRUE) %>% 
  column_spec(1) %>% 
  column_spec(2, width = "20em") %>% 
  column_spec(3, width = "20em")
```